Using Scagnostics to find Interesting Projections of Multivariate Data

### 2018-09-17 scagnostics.Rmd
Also comparing holes and cmass index.

```{r getProj}
getProj <- function(df, tPath, nameStr, size){
  sc <- tibble(convex=numeric(),
               clumpy=numeric(),
               skinny=numeric(),
               stringy=numeric(),
               dcor2d=numeric(),
               splines2d=numeric(),
               MIC=numeric(),
               TIC1=numeric(),
               t=numeric(),
               name=character(),
               size=numeric())
  n <- length(tPath)
  df <- rescale(df)
  for (i in 1:n) {
    dprj <- df %*% tPath[[i]]
    scagRes <- scagnostics(dprj)
    dcorRes <- dcor2d(dprj[,1], dprj[,2])
    splineRes <- splines2d(dprj[,1], dprj[,2])
    mineRes <- mine(dprj[,1], dprj[,2])
    sc <- add_row(sc, convex=scagRes["Convex"],
                  clumpy=scagRes["Clumpy"], skinny=scagRes["Skinny"],
                  stringy=scagRes["Stringy"], dcor2d=dcorRes,
                  splines2d=splineRes, MIC=mineRes$MIC,
                  TIC1=mineRes$TIC, t=i, name=nameStr, size=size)
  }
  maxTIC <- max(sc$TIC1)
  sc <- sc %>%
    mutate(TIC = TIC1/maxTIC) %>%
    select(-TIC1)

@@

```{r grandtourpath}
set.seed(1988)
t1 <- save_history(sphere100, grand_tour(2), max=4)
t1full <- as.list(interpolate(t1))
fullRes <- getProj(sin100, t1full, "sin", 100) %>%
  rbind(getProj(sin1000, t1full, "sin", 1000)) %>%
  rbind(getProj(sphere100, t1full, "sphere", 100)) %>%
  rbind(getProj(sphere1000, t1full, "sphere", 1000)) %>%
  rbind(getProj(pipe100, t1full, "pipe", 100)) %>%
  rbind(getProj(pipe1000, t1full, "pipe", 1000))
```

@@

## Transition
We repeat the same exercise, but now we are interested in how the index values change when going from the view in V1 vs V2 to the view in V5 vs V6.
```{r plannedtourpath}
m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
m2 <- matrix(c(0,0,0,0,1,0,0,0,0,0,0,1),ncol=2)
#this is silly but seems that first two entries are being ignored, so need some fake entries
m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
t2 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,m2)))
t2full <- as.list(interpolate(t2))
fullResPlanned <- getProj(sin100, t2full, "sin", 100) %>%
  rbind(getProj(sin1000, t2full, "sin", 1000)) %>%
  rbind(getProj(sphere100, t2full, "sphere", 100)) %>%
  rbind(getProj(sphere1000, t2full, "sphere", 1000)) %>%
  rbind(getProj(pipe100, t2full, "pipe", 100)) %>%
  rbind(getProj(pipe1000, t2full, "pipe", 1000))
```

```{r plannedgrandtour, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size"))
ggplot(fullResMeltPlanned, aes(x=t, y=value)) +
  geom_line(aes(color=name, linetype=factor(size))) + 
  facet_wrap(~variable, ncol=1, scales = "free_y")
```

>>>>

```{r getProj}
getProj <- function(df, tPath, nameStr, size){
  sc <- tibble(
    holes=numeric(),
    cmass=numeric(),
    convex=numeric(),
    clumpy=numeric(),
    skinny=numeric(),
    stringy=numeric(),
    dcor2d=numeric(),
    splines2d=numeric(),
    MIC=numeric(),
    TIC1=numeric(),
    t=numeric(),
    name=character(),
    size=numeric())
  n <- length(tPath)
  df <- rescale(df)
  for (i in 1:n) {
    dprj <- df %*% tPath[[i]]
    scagRes <- scagnostics(dprj)
    dcorRes <- dcor2d(dprj[,1], dprj[,2])
    splineRes <- splines2d(dprj[,1], dprj[,2])
    mineRes <- mine(dprj[,1], dprj[,2])
    holesRes <- holes(dprj)
    cmassRes <- cmass(dprj)
    sc <- add_row(sc, holes=holesRes, cmass=cmassRes,
                  convex=scagRes["Convex"],
                  clumpy=scagRes["Clumpy"], skinny=scagRes["Skinny"],
                  stringy=scagRes["Stringy"], dcor2d=dcorRes,
                  splines2d=splineRes, MIC=mineRes$MIC,
                  TIC1=mineRes$TIC, t=i, name=nameStr, size=size)
  }
  maxTIC <- max(sc$TIC1)
  sc <- sc %>%
    mutate(TIC = TIC1/maxTIC) %>%
    select(-TIC1)

@@

```{r grandtourpath}
set.seed(1988)
t1 <- save_history(sphere100, grand_tour(2), max=4)
t1full <- as.list(interpolate(t1))
fullRes <- getProj(sin100, t1full, "sin", 100) %>%
  rbind(getProj(sin1000, t1full, "sin", 1000)) %>%
  rbind(getProj(sphere100, t1full, "sphere", 100)) %>%
  rbind(getProj(sphere1000, t1full, "sphere", 1000)) %>%
  rbind(getProj(pipe100, t1full, "pipe", 100)) %>%
  rbind(getProj(pipe1000, t1full, "pipe", 1000))

@@

## Transition
We repeat the same exercise, but now we are interested in how the index values change when going from the view in V1 vs V2 to the view in V5 vs V6.
```{r plannedtourpath}
m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
m2 <- matrix(c(0,0,0,0,1,0,0,0,0,0,0,1),ncol=2)
#this is silly but seems that first two entries are being ignored, so need some fake entries
m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
t2 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,m2)))
t2full <- as.list(interpolate(t2))
fullResPlanned <- getProj(sin100, t2full, "sin", 100) %>%
  rbind(getProj(sin1000, t2full, "sin", 1000)) %>%
  rbind(getProj(sphere100, t2full, "sphere", 100)) %>%
  rbind(getProj(sphere1000, t2full, "sphere", 1000)) %>%
  rbind(getProj(pipe100, t2full, "pipe", 100)) %>%
  rbind(getProj(pipe1000, t2full, "pipe", 1000))

```{r plannedgrandtour, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size"))
ggplot(fullResMeltPlanned, aes(x=t, y=value)) +
  geom_line(aes(color=name, linetype=factor(size))) + 
  facet_wrap(~variable, ncol=1, scales = "free_y")
